---
title: "R and Data Science Introduction"
author: "Siddharth Jain"
date: "13 February 2019"
output: rmarkdown::github_document
---

## Introduction
**Data Science = Statistics/Maths + Coding + Domain/Business Knowledge**

Art of uncovering insights and trends behind data

## Data
3 main categories

* Structured (Databases)
* Semi Structured (XML/ JSON/ Web pages)
* Unstructured (Audio/ Video/ Image/ Natural Language)

Structured data = 20% of all data

## Big Data
Data whose scale, diversity, distribution and/or timeliness requires advanced technical architectures and analytics to enable insights.

Example: NY Stock Exchange

* Huge volume
* Complexity of datatypes and structures
* Speed of new data creation
* High growth

Drivers of Big Data:

* Opportunity to enable innovative new business models
* Potential for new insights
* Exponential data growth
* New formats of data
* New sources of data (live clicks, streaming data, social networks)
* Traditional solutions aren't working as well as they should
* Cost of data systems is increasing
* Open source data

**3Vs: Volume, Variety, Velocity**

**Importance** of Big Data:

* Cost reduction
* Time reduction
* Smart decision making
* Optimized offerings

## Data Engineering
Data engineers help Data scientists do their jobs more effectively.

Gather, collect, store, batch process, and serve data to data scientists.

* Wrangling: Manipulating raw data to make it useful for analytics/ for a ML model
* Cleaning: Missing values, insufficient parameters
* Preparation: Normalization, transforming input features, etc.

## R Programming

Some important Data Structures in R:

1. Vector (similar types of data)
2. Matrix (2 or more dimensional)
3. List (data of mixed types)
4. Data Frames (each component of same length)
5. Factors (Used to store categorical data)

#### Vectors
Use the ```c``` keyword to concatenate items into a vector.

```<-``` is the assignment operator in R. (```=``` can also be used in most cases)
```{r}
a <- c(1.8, 4.5, 3.3, 0.3)
print(typeof(a))
```

#### Matrix ,list
Creating a 3x3 matrix in R
```{r}
vector1 <- c(2,9,3)
vector2 <- c(10,16,17,13,11,15)

colnames = c("Col1", "Col2", "Col3")
rownames = c("Row1", "Row2", "Row3")
result <- array(c(vector1,vector2),dim = c(3,3), dimnames = list(colnames,rownames))

print(result)
```
