---
title: "R and Data Science Introduction"
author: "Siddharth Jain"
date: "13 February 2019"
output: rmarkdown::github_document
---

## Introduction
**Data Science = Statistics/Maths + Coding + Domain/Business Knowledge**

Art of uncovering insights and trends behind data

## Data
3 main categories

* Structured (Databases)
* Semi Structured (XML/ JSON/ Web pages)
* Unstructured (Audio/ Video/ Image/ Natural Language)

Structured data = 20% of all data

## Big Data
Data whose scale, diversity, distribution and/or timeliness requires advanced technical architectures and analytics to enable insights.

Example: NY Stock Exchange

* Huge volume
* Complexity of datatypes and structures
* Speed of new data creation
* High growth

Drivers of Big Data:

* Opportunity to enable innovative new business models
* Potential for new insights
* Exponential data growth
* New formats of data
* New sources of data (live clicks, streaming data, social networks)
* Traditional solutions aren't working as well as they should
* Cost of data systems is increasing
* Open source data

**3Vs: Volume, Variety, Velocity**

**Importance** of Big Data:

* Cost reduction
* Time reduction
* Smart decision making
* Optimized offerings

## Data Engineering
Data engineers help Data scientists do their jobs more effectively.

Gather, collect, store, batch process, and serve data to data scientists.

* Wrangling: Manipulating raw data to make it useful for analytics/ for a ML model
* Cleaning: Missing values, insufficient parameters
* Preparation: Normalization, transforming input features, etc.

## R Programming

Some important Data Structures in R:

1. Vector (similar types of data)
2. Matrix (2 dimensional)
3. List (data of mixed types)
4. Array (vector with one or more dimensions)
4. Data Frames (each component of same length)
5. Factors (Used to store categorical data)

#### R Data Types

* Logical (TRUE, FALSE)
* Numeric (12.3, 5, 1)
* Integer (2L, 34L, 0L)
* Complex (3+2i)
* Character ('a', "Hello")
* Raw (Use charToRaw to convert to ASCII values)

Everything in R is an object, and objects have attributes (identifiers).

* Name
* Dimension names (```dimnames```)
* Class
* Length

#### Vectors
Use the ```c``` keyword to concatenate items into a vector.

```<-``` is the assignment operator in R. (```=``` can also be used in most cases)
```{r}
a <- c(1.8, 4.5, 3.3, 0.3)
print(typeof(a))
```
Creating a random vector with 5 elements, printing its sum
```{r}
x <- rnorm(5)
print(x)
print(sum(x))
```
Other built in arithmetic functions

* sum() for total sum of vector
* var() for variation
* mean() for mean of the vector
* min() and max() to find minimum and maximum elements

Sequences in R
```{r}
x <- seq(1, 5)
print(x)
```
```by``` attribute can be passed to seq function to create steps
```{r}
y <- seq(-1, 0, by=0.1)
print(y)
```

#### Matrices
Creating a 3x3 matrix in R
```{r}
vector1 <- c(2,9,3)
vector2 <- c(10,16,17,13,11,15)

colnames = c("Col1", "Col2", "Col3")
rownames = c("Row1", "Row2", "Row3")
result <- array(c(vector1,vector2),dim = c(3,3), dimnames = list(colnames,rownames))

print(result)
```
Creating an empty 2x3 matrix
```{r}
m = matrix(0.0, nrow=2, ncol=3) # 2x3
print(m)
```
Checking the class of the object
```{r}
print(class(m))
```
Checking type of values
```{r}
print(typeof(m))
```
Checking length
```{r}
print(length(m))
```

#### Lists
Collections of mixed type data objects

The objects in a list are known as its components

Example
```{r}
lst <- list(name = "Sid", height = "6", num_children = 3, child_ages = c(6, 7, 9))
print(lst[[1]])
print(lst$height)
```
To print 1st component of child_ages
```{r}
print(lst[[4]][1])
```

#### Data Frames
Similar to matrix, may contain categorical as well as numerical data.

It is like a list with components as columns of a table.

```{r}
measrmts <- data.frame(gender = c("F", "M", "M"), ht = c(180, 155, 165), wt = c(60, 80, 75))
print(measrmts)
print(measrmts[1, 2])
```
Character vectors in data frame are stored as factor
```{r}
class(measrmts$gender)
```







