---
title: "R Analysis of Titanic Dataset"
author: "Siddharth Jain"
date: "02 March 2019"
output: rmarkdown::github_document
---

## Introduction
We are analysing the various factors determining survival rates on the Titanic disaster. To do this, we will use various different features available about the passengers and visualize them.

### Importing the data
```{r}
library("naniar")
library("ggplot2"); theme_set(theme_minimal())
library("dplyr")
library("tidyr")
test = read.csv('test.csv', stringsAsFactors = FALSE)
train = read.csv('train.csv', stringsAsFactors = FALSE)
full <- bind_rows(train,test)
LT=dim(train)[1]
head(LT)
```

```{r}
str(full)
```

### Visualizing missing data in the dataset
```{r}
vis_miss(train)
```

```{r}
missing_values <- full %>% summarize_all(funs(sum(is.na(.))/n()))
missing_values <- gather(missing_values, key="feature", value="MissingPercentage")

missing_values %>% 

  ggplot(aes(x=reorder(feature,-MissingPercentage),y=MissingPercentage)) +

  geom_bar(stat="identity",fill="red")+

  coord_flip()+theme_bw()
```

Age has many missing values.


### Taking care of missing Age values
Replacing the missing values with the mean age of all the passengers on the Titanic
```{r}
full <- full %>%

    mutate(

      Age = ifelse(is.na(Age), mean(full$Age, na.rm=TRUE), Age),

      `Age Group` = case_when(Age < 13 ~ "Age.00.12", 

                                 Age >= 13 & Age < 18 ~ "Age.13.17",

                                 Age >= 18 & Age < 60 ~ "Age.18.59",

                                 Age >= 60 ~ "Age.60.Ov"))
```

### Relationship between sex and survival, and subsequently between age and survival
```{r}
ggplot(data=full[1:LT,],aes(x=Sex,fill=Survived))+geom_bar()

# First we'll look at the relationship between age & survival
ggplot(full[1:891,], aes(Age, fill = factor(Survived))) + 
  geom_histogram() + 
  # Including gender since we know it's important
  facet_grid(.~Sex) + 
  theme_minimal()
```

### Survival rate based on family size
```{r}
full$Fsize <- full$SibSp + full$Parch + 1
full$Family <- paste(full$Surname, full$Fsize, sep='_')
# Use ggplot2 to visualize the relationship between family size & survival
ggplot(full[1:891,], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') +
  theme_minimal()
```

We can see that single people and families with size>4 had lower survival rates.


### Making a random forest model
```{r}
library('randomForest')
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + 
                                            Fare + Embarked,
                                            data = train, na.action = na.roughfix)

# Show model error
plot(rf_model, ylim=c(0,0.36))
legend('topright', colnames(rf_model$err.rate), col=1:3, fill=1:3)
```